---
output: 
  html_document: 
    theme: flatly
    toc: no
---
World Food Facts by Christian Burkhart
========================================================


The **World Food Facts** dataset is a french initiative. It tries to collect food products from all over the world. The dataset is provided by **Open Food Facts** under an Open Database License (http://opendatacommons.org/licenses/odbl/1.0/). Customers and food manufacturers can freely add products.

In this Exploratory Data Analysis I will try to get some insights into the food products of Germany. Being from Germany I am interested in the nutritional value of the products found in the German supermarket.


```{r Libraries, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(reshape2)
```

```{r Constants, echo=FALSE}
COLOR_FILL<- "#CD6771"
COLOR_TWO <- "#0D5047"
COLOR_BORDER <- "#F6A4AC"
```

```{r Helper functions, echo=FALSE}
histogram_helper <- function(variable, xlab, bwidth) {
  # Renders a histogram of a specific variable 
  #
  # Args:
  #  variable: The variable that should be displayed.
  #  xlab: The title of the x axis
  #  bwidth: The width of the bins for the histogram.
  #
  # Returns:
  #  A ggplot2 histogram
  histogram <- ggplot(food_data_germany, aes(variable)) +
    geom_histogram(binwidth = bwidth, fill = COLOR_FILL, colour = COLOR_BORDER) +
    labs(x = xlab)
  return(histogram)
}

scatterplot_ingredients <- function(xvariable, yvariable, xl, yl) {
  # Renders a scatterplot for different food ingredients.
  #
  # Args:
  #  xvariable: The variable for the x-axis.
  #  yvariable: The variable for the y-axis.
  #  xlab: String for the x-axis.
  #  ylab: String for the y-axis.
  #
  # Returns:
  #  ggplot2 scatterplot.
  
  scat <- ggplot(NULL, aes(x = xvariable, y = yvariable)) +
    geom_point(color = COLOR_FILL) +
    xlab(xl) +
    ylab(yl)
    #labs(xlab = xl, ylab = yl)

  return(scat)
}

density_plot <- function(xvariable, fill, xl) {
  # Renders a density plots of a variable with different
  # factors
  #
  # Args:
  #  xvariable: The variable for the x-axis.
  #  fill: A factor variable.
  #  xlab: String for the x-axis.
  #
  # Returns:
  #  Density plots
  
  ggplot(food_common, aes(x = xvariable, fill = fill)) +
    geom_density(alpha = 0.4) +
    xlab(xl) +
    ylab("density")
}
```


```{r Load data, echo=FALSE, cache=TRUE}
setwd("~/Repositories/Udacity_Explore-and-Summarize-Data")
food_data <- read.csv(file.path("data", "FoodFacts.csv"), header = TRUE, sep = ",", stringsAsFactors = TRUE)
food_data_germany <- filter(food_data, countries_en == "Germany")
```


# Data Cleaning
There are in total `r nrow(food_data)` food products in the dataset. Each food product has `r ncol(food_data)` variables. Not every variable will be of interest for this analysis. I will focus my analysis on products from **products from Germany**. When I filter the database for German products the entire database is reduced `r nrow(food_data_germany)` products. In order to select variables that contain meaningful information we have to filter the dataset for NAs. Variables with a lot of **NAs** should not be included in the analysis. For this analysis I will remove features with more than 40% of NAs.

```{r NAs, echo=FALSE, warning=FALSE}
# Code taken and accustomed from
# https://www.kaggle.com/jelaurent/d/openfoodfacts/world-food-facts/openfoodfacts-exploring-the-dataset/notebook
(na_variables <- food_data_germany %>%
  melt(measure.vars = 1:159) %>%
  group_by(variable) %>% 
  summarize(perc_na = round(sum(is.na(value))/length(value), digits = 3)) %>%
  arrange(desc(perc_na)) %>%
  filter(perc_na < .4) %>%
  as.data.frame)
```

Most features do contain a lot of missing values. Out of `r ncol(food_data_germany)`, `r nrow(na_variables)` meet the NA criteria (less than 40%). Not all of these features are of interest.

```{r Filter data frame, echo=FALSE, warning=FALSE, cache=TRUE}
food_germany <- select(food_data_germany, product_name, generic_name,
                       brands, brands_tags, categories_en, origins, pnns_groups_1, 
                       pnns_groups_2, manufacturing_places, additives_en,
                       nutrition_grade_fr, main_category_en, energy_100g, 
                       fat_100g, saturated_fat_100g, carbohydrates_100g, 
                       sugars_100g, proteins_100g, salt_100g, 
                       nutrition_score_fr_100g, nutrition_score_uk_100g)
glimpse(food_germany)
```

```{r Colnumber of important features, echo=FALSE}
brands_col <- 3
pnns_1_col <- 7
pnns_2_col <- 8
fat_col <- 14
saturated_fat_col <- 15
carbohydrates_col <- 16
proteins_col <- 18
nutrition_score_fr_col <- 20
nutrition_score_uk_col <- 21
```

We managed to drill down the database to `r ncol(food_germany)` variables. These features now contain less than 40% of NAs and can be used for further analysis.

```{r Tidy pnss features, echo=FALSE}
# Borrowed idea from jlaurent:
# https://www.kaggle.com/jelaurent/d/openfoodfacts/world-food-facts/openfoodfacts-exploring-the-dataset/notebook

# Capitalize first letter of feature
food_germany$pnns_groups_1 <- gsub("^([a-z])(\\w+)", "\\U\\1\\L\\2", food_germany$pnns_groups_1, perl = TRUE)
food_germany$pnns_groups_2 <- gsub("^([a-z])(\\w+)", "\\U\\1\\L\\2", food_germany$pnns_groups_2, perl = TRUE)

# Replace 'unkown' with empty string
food_germany$pnns_groups_1 <- gsub("unknown", "", food_germany$pnns_groups_1)

# Replace dash with space
food_germany$pnns_groups_1 <- as.factor(gsub("-", " ", food_germany$pnns_groups_1))

# Lowercase variables
food_germany$pnns_groups_1 <- gsub(pattern = '([[:upper:]])', perl = TRUE, replacement = '\\L\\1', food_germany$pnns_groups_1)
food_germany$pnns_groups_2 <- gsub(pattern = '([[:upper:]])', perl = TRUE, replacement = '\\L\\1', food_germany$pnns_groups_2)
```

# Univariate Plots Section

## Macronutritions
I will begin to explore the **macronutritions**: That is, **fat**, **proteins**, and **carbohydrates**. 

```{r Histogram fat, echo=FALSE}
histogram_helper(food_germany$fat_100g, "fat_100g", bwidth = 2)
```

Most products do not contain a lot of **fat**. The distribution is **skewed left**.

For the purpose of a healthy diet is is crucial to investigate **saturated fat** as it has been linked to *cardiovascular diseases*.

```{r Histogram saturated fat, echo=FALSE}
histogram_helper(food_germany$saturated_fat_100g, "saturated_fat_100g", bwidth = 2)
```

Most products do not contain a lot of **saturated fat**. The distribution is sort of bimodal. There is tiny peak of products that contain around 50g of saturated fat per 100g.

```{r Products with ~ 50g of saturated fat per 100g, echo=FALSE}
filter(food_germany, saturated_fat_100g > 49, saturated_fat_100g < 60)$product_name
```

These are **butter** products. In terms of a healthy diet one shouldn't eat too much butter.

```{r Histogram proteins, echo=FALSE}
histogram_helper(food_germany$proteins_100g, "proteins_100g", bwidth = 2)
```

The distribution of **protein** is **skewed to the left**. There are few products with a lot of protein. These are **food supplements** for weight loss or muscle training. 

```{r Products high in protein, echo=FALSE}
filter(food_germany, proteins_100g > 75)$product_name
```

```{r Histogram carbohydrates, echo=FALSE}
histogram_helper(food_germany$carbohydrates_100g, "carbohydrates_100g", bwidth = 2)
```

The distribution of **carbohydrates** is also **skewed left**. Most products do not contain a lot of carbohydrates. The distribution is slightly **bimodal**.

```{r Histogram sugar, echo=FALSE}
histogram_helper(food_germany$sugars_100g, "sugars_100g", bwidth = 2)
```

The distribution of sugar is sort of similar. It is also **skewed left** and **bimodal**. 

## Nutritional Value

### Nutritional score
The database has two features with information about the product's nutritional value. There is a score from France (**nutrition_score_fr_100g**) and a score from the UK (**nutrition_score_uk_100g**). Let us see how they differ.

```{r Histogram nutritional scores, echo=FALSE}
# Melt data frame for nutritional score features
nutrition_score_dataframe <- food_germany %>%
  melt(measure.vars = nutrition_score_fr_col:nutrition_score_uk_col)

## Histograms of nutritional score features
ggplot(nutrition_score_dataframe, aes(x = value)) +
  geom_histogram(binwidth = 2, fill = COLOR_FILL, colour = COLOR_BORDER) +
    labs(x = "Nutrional Value") +
    facet_wrap(~ variable)

# Density plot 
ggplot(nutrition_score_dataframe, aes(x = value, fill = variable)) +
  geom_density(alpha = .3) +
  labs(x = "Nutritional Value")
```

Obviously the variables are not drastically different from another. The British system seems to be more conservative (the lower the value the healthier). For further analysis I will stick to the **French** system. 

From a first glimpse at the nutritional value, the distribution looks a little bit **bimodal**. There is a high peak at around a score of 0. These might be water products that are neither nutritional valuable nor non-valuable. Another peak seems to be around a score of 15. 

## Nutritional grade

Apart from the nutritional score there is a feature about the nutrional grade of the products (**nutrition_grade_fr**). The grade system is very similar to the grade system in schools. **a** is supposed to be a healthy product, whereas **e** is an unhealthy product.

```{r Nutritional grade of products, echo=FALSE}
# Barchart of nutritional grades
ggplot(food_germany, aes(x = nutrition_grade_fr)) +
  geom_bar(fill = COLOR_FILL)
```

Many products do not have a grade. The graded products have a sort of uniform distribution. 

## Other variables

```{r Histogram of kcal, echo=FALSE}
food_germany$kcal_100g <- food_germany$energy_100g * 0.239006
histogram_helper(food_germany$kcal_100g, "kcal_100g", bwidth = 10)
```

Energy in food is most often measured in **kilojoule**. Customers are usually more accustomed to **kilocalories**. Therefore I created a new variable **kcal_100g** that used the conversion of the following website: http://www.rapidtables.com/convert/energy/kj-to-kcal.htm. 

The distribution has three peaks and is skewed left. The mean is `r mean(food_germany$kcal_100g, na.rm = TRUE)`. 

```{r Products high in kcal, echo=FALSE}
filter(food_germany, kcal_100g > 700)$product_name
```

Olive oil, sunflower oil and butter are very high in kcals. 

```{r Histogram salt, echo=FALSE}
histogram_helper(food_germany$salt_100g, "salt_100g", bwidth = 1)
```

There is not much **salt** in most products. Just very few have a lot of salt. I reckon these are salt products :) 

It would be also interesting to find out which brands produce most products?

```{r Barchart common brands, echo=FALSE}
common_brands_vector <- sort(table(food_germany$brands), decreasing = TRUE)[2:30]
common_brands_names <- names(common_brands_vector)
common_brands_df <- data.frame(brands = common_brands_names, frequency = common_brands_vector)
common_brands_df <- transform(common_brands_df, brands = reorder(brands, -frequency))

ggplot(common_brands_df, aes(x = brands, y = frequency)) + 
  geom_bar(stat = "identity", fill = COLOR_FILL, color = COLOR_BORDER) +
   theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

Most products sold in Germany are made by **Gut & Günstig**. **Alnatura** comes in second place. 

```{r Most common groups, echo=FALSE}
ggplot(food_germany, aes(pnns_groups_1)) +
  geom_bar(fill = COLOR_FILL)
```

Most products are **dairy products**. **Beverages** and **sugary snacks** come in second place. Almost half of the products do not a group assigned. 


# Univariate Analysis

### What is the structure of your dataset?
There are `r nrow(food_germany)` products with `r ncol(food_germany)` features in the dataset.

```{r Glimpse at dataset, echo=FALSE}
glimpse(food_germany)
```

The dataset contains factors and numerics. The factor's level differ heavily. There are features with just 6 levels and features with over 20,000 levels. 


### What is/are the main feature(s) of interest in your dataset?

The main features of the dataset are features that contain information that is relevant for a healthy diet. These are all **macronutritions**, **nutritional_score_fr_100g**, **nutritional_grade_fr** and **kcal_100g**.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

For further investigation I will look into the **pnns_groups_1**, **pnns_groups_2**, and **brands** features. It will be interesting to find out if some types of products are better of worse for a healthy diet. It might also be interesting to check whether some brands are healthier than others.

### Did you create any new variables from existing variables in the dataset?
I created a new variable **kcal_100g**. I am more accustomed to kcal than kilojoule as an indicator of energy. In order to find the most common **brands** I created the dataframe **common_brands_df**. 

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?

The **pnns_groups_1** variable had to be tidied. In order to avoid duplicates, I set all entries to lowercase and replaced a dash with a blank space. I also capitalized the first letter of the feature. I filtered the whole data set for german entries. Being from Germany these are the products that interest me. 

Most distributions of **macronutritions** are skewed left. This is not an unusual pattern though as they add up to about 100g. What striked me was that there is group of products that are very high in protein. These are bodybuilding products. No other products contain so much protein per 100g. What also struck me was the high amount of saturated fat in butter. Butter seems to be especially unhealthy as it also contains a lot of kcals. Olive and sunflower oil are also very high in kcals. 


<!----------------------------- Bivarate Analysis ------------------------------------------------- -->
# Bivariate Plots Section
```{r Scatterplot matrix, echo=FALSE}
pairs(~ proteins_100g +
        fat_100g+carbohydrates_100g +
        saturated_fat_100g +
        nutrition_score_fr_100g +
        sugars_100g, 
      col = COLOR_FILL, pch = 1, cex = 0.2, data = food_germany,
      main="Simple Scatterplot Matrix")
```

```{r Correlation table, echo=FALSE}
round(cor(food_germany[, c(13:22)], use = "complete.obs"), 2)
```

**Energie** and **fat** are highly correlated (r = .81). This makes sense as one gram of fat contains 9 kcal and therefore more energy than carbohydrates and protein. **Sugar* and **saturated fat** is not correlated (r = .04). This indicates that there are no sugary products that contain a lot of saturated fat. 

Looking at the scatterplots there is an interesting pattern. Some products seem to contain a lot of protein and a tiny amount of carbohydrates and fat. 

```{r Scatterplot fat and protein, echo=FALSE}
scatterplot_ingredients(food_germany$fat_100g, food_germany$proteins_100g, 
                        "fat_100g", "proteins_100g")
```

When we plot **fat** and **protein** two interesting patterns emerge: (1) **Bodybuilding products** are high in protein but very low in fat. (2) Olive oils are on the other extreme. They are very high in fat but low in protein.

```{r high protein low fat products, echo=FALSE}
(high_protein_low_fat <- filter(food_germany, proteins_100g > 75, fat_100g < 20)$product_name)
```

```{r High in fat and low in protein, echo=FALSE}
(high_fat_low_protein <- filter(food_germany, proteins_100g < 5, fat_100g > 90)$product_name)
```

Let us have a look at **fat** and **carbohydrates**.

```{r Scatterplot fat and carbohydrates, echo=FALSE}
scatterplot_ingredients(food_germany$fat_100g, food_germany$carbohydrates_100g,
                        xl = "fat_100g", yl = "carbohydrates_100g")
```

There are no special food groups in this scatterplot. What is interesting is the **hole** inside the scatterplot. Very few products contain around 25% of both protein and fat.

What about **carbohydrates** and **protein**?

```{r Scatterplot carbohydrates and protein, echo=FALSE}
scatterplot_ingredients(food_germany$carbohydrates_100g, food_germany$proteins_100g,
                        "carbohydrates_100g", "proteins_100g")
```

Again we can find the same pattern. The **bodybuilding products** are very high in protein but low in carbohydrates. Apart from that there is nothing particularly special about the dataset. 

```{r Scatterplot nutrional value and fat, echo=FALSE}
scatterplot_ingredients(food_germany$nutrition_score_fr_100g, 
                        food_germany$saturated_fat_100g,
                        "nutrition_score_fr_100g",
                        "saturated_fat_100g")
```

**Saturated fact** is often associates with an unhealthy diet. **Coconut oil** and **coconut fat** are both high in saturated fat and nutritional score. 

```{r Products high in saturated fat and nutritional score, echo=FALSE}
(high_saturated_fat_nutritional_score <- filter(food_germany, nutrition_score_fr_100g > 18, saturated_fat_100g > 75)$product_name)
```

What about **saturated fat** and **fat**?

```{r Scatterplot saturated fat and fat, echo=FALSE}
scatterplot_ingredients(food_germany$fat_100g, food_germany$saturated_fat_100g,
                        "fat_100g", "saturated_fat_100g")
```

Most products contain few fat and and saturated fat. Two products seem to contain a high amount of saturated fat. 

```{r Products high in saturated fat, echo=FALSE}
(high_saturated_fat <- filter(food_germany, saturated_fat_100g > 75,
                              fat_100g > 95)$product_name)
```

Again, this is our **coconut oil** and **coconut fat**.


# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

Nutritional data all fall within a diagonal of a scatterplot. As the macronutritions features all add up to maximal 100g the features can not add up above 100g. 

Linear relationsships do not occur. Almost all relationships are positive. Most of these relationships are small. 

When I plotted fat and carbohydrates with protein some products cluster around a high protein level with a low level of the other two macronutritions. These are bodybuilding products as muscles need protein to grow. 

**Products high in saturated fat** usually have a low nutrition score. This is in line with the usually recommendation to avoid saturated fats. 

Most products that contain **fat** do not contain a high amount of **saturated fat**. Coconut oil is one exception. It even as a high nutrition score as is evident in the scatterplot of saturated fat and nutrition score. 

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

No.

### What was the strongest relationship you found?

One of the strongest relationship was between **fat** and **saturated fat** (r = . 81). The nutritional score features correlate strongly with another (r = .97). 


<!----------------------------- Multivariate Analysis ------------------------------------------------- -->
# Multivariate Plots Section

```{r Filter data frame for common brands, echo=FALSE}
food_types <- filter(food_germany, pnns_groups_1 %in% pnns_groups_1[! pnns_groups_1 %in% c("unknown", "")])
food_common <- filter(food_types, brands %in% common_brands_names[c(1:5)])
```

```{r Density plot carbohydrates food category, echo=FALSE}
density_plot(food_common$carbohydrates_100g, food_common$pnns_groups_1, "carbohydrates_100g")
```

Sugary snacks are heavy in carbohydrates. Beverages usually contain about 15 gram of carbohydrates.

```{r Density plot sugars food category, echo=FALSE}
density_plot(food_common$sugars_100g, food_common$pnns_groups_1, "sugars_100g")
```

The distribution for sugar is sort of similar. Beverages contain a specific amount of sugar. Salty snacks and sugary snacks contain most sugar. 

```{r Density plot proteins food category, echo=FALSE}
density_plot(food_common$proteins_100g, food_common$pnns_groups_1, "proteins_100g")
```

Milk and dairy products contain most protein. Fruits and vegetables are usually low in protein and also contain a specific amount.

```{r Density plot proteins nutrition_grade_fr, echo=FALSE}
density_plot(food_common$proteins_100g, food_common$nutrition_grade_fr, "proteins_100g")
```

Plotting the distribution of graded products no specific pattern emerges. The amount of protein does not determine wheater a product is considered healthy or not healthy. 

```{r Kilocalories of different food groups, echo=FALSE}
ggplot(food_types, aes(x = kcal_100g, fill = pnns_groups_1)) +
  geom_density(alpha = .5)
```

**Fat and sauces** are very heavy in kcals. Fish, meat, and eggs usually do not contain a lot kcals. 

```{r Nutritional score of most common brands, echo=FALSE}
ggplot(food_common, aes(x = nutrition_score_fr_100g, fill = brands)) +
    geom_density(alpha = 0.4) +
    labs(x = "nutrition_score_fr_100g")
```

Milkbona seems to be the most healthy brand. This is a biased representation because Milkbona sells milk products that usually are considered more healthy. Except **Edeka** every brand has a bimodal distribution. Many products can be considered healthy whereas the other half can be considered unhealthy. 

```{r Nutritional Score of beverages, echo=FALSE}
# Get rows containing beverages
beverage_rows_pnns_1 <- grep("beverages", food_germany$pnns_groups_1)
beverage_rows_pnns_2 <- grep("beverages", food_germany$pnns_groups_2)

# Create new data frame
beverages <- food_germany[unique(c(beverage_rows_pnns_1, beverage_rows_pnns_2)), ]

# Get alcoholic beverages
alcohol_beverages <- grep("alcohol", beverages$pnns_groups_2)

# Generate vector of alcoholic beverages
alcohol <- c()
for (i in c(1:nrow(beverages))) {
  if (i %in% alcohol_beverages) {
    alcohol <- c(alcohol, "alcoholic")
  } else {
    alcohol <- c(alcohol, "non-alcoholic")
  }
}

# Add vector to beverages
beverages$alcohol <- as.factor(alcohol)

# Plot beverages
ggplot(beverages, aes(x = nutrition_score_fr_100g)) +
  geom_density(fill = COLOR_FILL) +
  facet_grid(~ alcohol)
```

It turned out that alcoholic products do not contain data about the nutritional score. Non-alcoholic beverages show a bimodal distribution. Most drinks are low in nutritional score. Some products are considered unhealthy. Most of these products are lemonades: 

```{r Unhealthy drinks, echo=FALSE}
(filter(beverages, nutrition_score_fr_100g > 10)$product_name)
```


```{r Nutrional score of product types, echo=FALSE}
# Filter products types for meaningful types
common_types <- filter(food_germany, pnns_groups_1 %in% pnns_groups_1[! pnns_groups_1 %in% c("unknown", "")])

# Plot most common
ggplot(common_types, aes(x = nutrition_score_fr_100g)) +
  geom_histogram(binwidth = 2, fill = COLOR_FILL, colour = COLOR_BORDER) +
    labs(x = "Nutrional Value") +
    facet_wrap(~ pnns_groups_1)
```

**Sugary** snacks obviously are not very healthy. Appart from sugary snacks there is no product type that is of bad nutritional value per se. **Cereals and potatoes** seem to be beneficial for health. The same goes for **Fruits and vegetables**. **Milk and dairy products** have a bimodal distribution. This indicates that half of these products are not very healthy. **Beverages** by and large seem to be reasonably good. The peak of the distribution might be because of **water products**. 

```{r Boxplots of food types, echo=FALSE}
ggplot(common_types, aes(x = pnns_groups_1, nutrition_score_fr_100g)) +
  geom_boxplot(fill = COLOR_FILL) +
  facet_wrap(~ nutrition_grade_fr) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

This plot shows the relationship between the nutritional grade and score variable. They correlate positively. Out of the food types that get a grade **a** fruits and vegetables are the most healthy.



# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

The data strengthened the common fact that **sugary snacks** and **salty snacks** are considered unhealthy. They not only contain a lot of kcals, they are also very high in carbohydrates (especially sugars). 

### Were there any interesting or surprising interactions between features?

What I found interesting is that beverages usually have a specific amount of carbohydrates and sugars. Usually they contain between 5 and 30 grams of carbohydrates. A lot of patterns weren't quite surprising. **Sugary snacks** are high in sugar and considered unhealthy. What surprised me is that beverages do not contain a lot of kcals. 

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r , echo=FALSE}
ggplot(common_types, aes(x = pnns_groups_1, nutrition_score_fr_100g)) +
  geom_boxplot(fill = COLOR_FILL) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Boxplot")
```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection
