World Food Facts by Christian Burkhart
========================================================


The **World Food Facts** dataset is a french initiative of a collection of food products around the world. The dataset is provided by **Open Food Facts** under an Open Database License (http://opendatacommons.org/licenses/odbl/1.0/). Customers and food manufacturar can freely add products. Every row within in the database represents a food product. 

In this Exploratory Data Analysis I will try to get some insights into interesting patterns of food that is eaten around the world. 


```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
library(dplyr)
```

```{r echo=FALSE, cache=TRUE, Load_the_Data}
setwd("~/Repositories/Udacity_Explore-and-Summarize-Data")
food_data <- read.csv(file.path("data", "FoodFacts.csv"), header = TRUE, sep = ",", stringsAsFactors = TRUE)
```

# Univariate Plots Section
```{r echo=FALSE, Univariate_Plots}

```

# Univariate Analysis

### What is the structure of your dataset?

```{r, echo=FALSE}
dim(food_data)
```

There are in total 65503 food products in the dataset. Each food product has 159 variables. Not every variable will be of interest for this analysis. Therefore I will select the dataset for variables that seem to be of interest. 

```{r, cache=TRUE}
food_data_selection <- select(food_data, product_name, generic_name, quantity,
                              brands, categories, categories_en, origins, 
                              origins_tags, manufacturing_places, countries_en,
                              serving_size, additives_en, nutrition_grade_uk,
                              nutrition_grade_fr, pnns_groups_1, pnns_groups_2,
                              states_en, main_category_en, energy_100g,
                              energy_from_fat_100g, fat_100g, saturated_fat_100g,
                              monounsaturated_fat_100g, polyunsaturated_fat_100g, 
                              omega_3_fat_100g, omega_6_fat_100g, omega_9_fat_100g,
                              trans_fat_100g, cholesterol_100g, carbohydrates_100g,
                              sugars_100g, sucrose_100g, glucose_100g, fructose_100g,
                              lactose_100g, maltose_100g, fiber_100g, proteins_100g,
                              salt_100g, alcohol_100g, vitamin_a_100g, 
                              vitamin_d_100g, vitamin_e_100g, vitamin_k_100g, 
                              vitamin_c_100g, vitamin_b1_100g, vitamin_b2_100g, 
                              vitamin_b6_100g, vitamin_b9_100g, vitamin_b12_100g,
                              caffeine_100g, nutrition_score_fr_100g, 
                              nutrition_score_uk_100g)
```

```{r, echo=FALSE}
str(food_data_selection)
```

The selected dataset now contains  variables. For further analysis it would be good to save the columns that can be grouped together.

```{r}
fat_columns <- c(22:28)
omega_fat_columns <- c(25:27)
vitamin_columns <- c(41:50)
```

I will focus my analysis on products from Germany. Living in Germany I am interested in the products that I find in the supermarket. 

```{r}
food_data_germany <- filter(food_data_selection, countries_en == "Germany")
```

There are  products in this dataset.

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?



# Bivariate Plots Section
```{r echo=FALSE, Bivariate_Plots}

```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

### What was the strongest relationship you found?




# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection
